Effect of Beam Chromaticity on Foregrounds in Wide-Field Measurements of Redshifted 21~cm Power Spectra

### 2016-03-05 ForegroundSimulations/ms.tex
Add Abstract, and minor edits
From f3ce757 (fgsim):
====================

In ms.tex,

(*) Add Abstract
(*) Update caption of figure 6
(*) Minor edits throughout

@@

The redshifted neutral hydrogen from the IGM in this epoch has been identified to be one of the most promising and direct probes of the EoR \citep{sun72,sco90,mad97,toz00,ili02}. Numerous experiments using low frequency radio telescopes targeting the redshifted 21~cm line from the spin-flip transition of H{\sc i} have become operational such as the Murchison Widefield Array \citep[MWA;][]{lon09,bow13,tin13}, the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, the Low Frequency Array \citep[LOFAR;][]{van13} and the Giant Metrewave Radio Telescope EoR experiment \citep[GMRT;][]{pac13}. These instruments have sufficient sensitivity for a statistical detection of the EoR signal via estimating the spatial power spectrum of the redshifted H{\sc i} temperature fluctuations \citep{bea13,thy13}. These instruments are intended to be precursors and pathfinders to the next generation of low frequency radio observatories such as the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} \citep[HERA;][]{deb16} and the Square Kilometre Array\footnote{\url{https://www.skatelescope.org/}} (SKA). These next-generation instruments will advance the capability from a mere statistical detection of the signal to a direct three-dimensional tomographic imaging of the H{\sc i} during the EoR.

@@

We note that these visibilities and delay power spectra are dependent on right ascension (RA) of the pointing as well as the center of the frequency band. We defer expressing visibilities and delay power spectra explicitly as a function of pointing or center frequency (or redshift) to later sections where it is necessitated.

@@

@@

Considering that the dynamic range required to suppress foreground contamination in the {\it EoR window} may probably be even higher than that provided by a {\it Blackman-Harris} window function, we use a modified version given by the convolution of a {\it Blackman-Harris} window with itself:
\begin{align}\label{eqn:bhw2}
  W(f) &= W_\textrm{BH}(f) \ast W_\textrm{BH}(f).
\end{align}
In its Fourier domain, this window has a response that is obtained by squaring of the {\it Blackman-Harris} window response and thus improves the dynamic range further in the power response by another 100 -- 120~dB while $\epsilon\approx 42\%$. This is an enormous gain in sidelobe characteristics for a negligible loss in over sensitivity relative to a simple {\it Blackman-Harris} weighting. This will ensure that the contamination resulting from spillover of foregrounds along the line-of-sight $k$-modes are limited by intrinsic spectral structures in the foregrounds and the beam patterns and not by sidelobes from spectral weighting. In this paper, we apply this modified spectral weighting function defined in equation~\ref{eqn:bhw2} whenever foregrounds are represented in the Fourier delay domain.

@@

\begin{figure*}[htb]
\centering
\subfloat[][EoR model 1, 150~MHz subband ]{\label{fig:21cmfast-fg-ratio-150}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_21cmfast_chrmbeam_sim_150.0_MHz_subband.eps}}
\subfloat[][EoR model 2, 150~MHz subband ]{\label{fig:lidz-fg-ratio-150}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_lidz_chrmbeam_sim_150.0_MHz_subband.eps}} \\
\subfloat[][EoR model 1, 170~MHz subband ]{\label{fig:21cmfast-fg-ratio-170}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_21cmfast_chrmbeam_sim_170.0_MHz_subband.eps}}
\subfloat[][EoR model 2, 170~MHz subband ]{\label{fig:lidz-fg-ratio-170}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_lidz_chrmbeam_sim_170.0_MHz_subband.eps}}
\caption{Lower limits on EoR sensitivity, $\rho(\boldsymbol{b},f_0,\alpha)$, in {\it foreground avoidance} modes for different HERA-19 baselines as a function of RA ($\alpha$) of pointing on the sky. Each panel corresponds to the EoR model and subband indicated. The baseline vector is specified by its length (in m) and orientation (in degrees) at the bottom and top of the $x$-axis respectively. The color scale is logarithmic. Even with worst-case sensitivity, clearly HERA will be able detect EoR drawn from independent models on all baselines for a large fraction of a full 24~hour sky transit cycle.}
\label{fig:eor-fg-ratios}
\end{figure*}

>>>>

Unaccounted systematics from foregrounds and instruments prevent present day experiments from detecting redshifted 21~cm signal from the Epoch of Reionization (EoR). In-depth understanding of these effects is critical for the success of such experiments. We identify spectral structures (chromaticity) in the antenna power pattern as one such systematic that significantly affects all EoR measurements. Increase in chromaticity of the antenna beam causes the already bright foreground power to leak well beyond ideal limits and contaminate the otherwise clean EoR signal modes. A primary source of this chromaticity is reflections in the antenna-feed assembly and between structures in neighboring antennas. Using precise foreground simulations taking wide-field effects into account, we provide a framework to set cosmologically-motivated upper limits on these reflections to prevent further EoR signal degradation. We use the Hydrogen Epoch of Reionization Array (HERA) as an example. Using a simple and yet high dynamic range spectral weighting scheme, we demonstrate that EoR detection using HERA will not be impeded by such spectral systematics even in a conservative scenario that does not require removal of foregrounds. We also show that all baselines in a 19-element HERA layout are capable of detecting EoR over a substantial observing window on the sky.

@@

The redshifted neutral hydrogen from the IGM in this epoch has been identified to be one of the most promising and direct probes of the EoR \citep{sun72,sco90,mad97,toz00,ili02}. Numerous experiments using low frequency radio telescopes targeting the redshifted 21~cm line from the spin-flip transition of H{\sc i} have become operational such as the Murchison Widefield Array \citep[MWA;][]{lon09,bow13,tin13}, the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, the Low Frequency Array \citep[LOFAR;][]{van13} and the Giant Metrewave Radio Telescope EoR experiment \citep[GMRT;][]{pac13}. These instruments have sufficient sensitivity for a statistical detection of the EoR signal via estimating the spatial power spectrum of the redshifted H{\sc i} temperature fluctuations \citep{bea13,thy13}. These instruments are intended to be precursors and pathfinders to the next generation of low frequency radio observatories such as the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} \citep[HERA;][]{deb16} and the Square Kilometre Array\footnote{\url{https://www.skatelescope.org/}} (SKA). These next-generation instruments will advance the capability from a mere statistical detection of the signal to a direct three-dimensional tomographic imaging of the H{\sc i} from the EoR.

@@

@@

We note that these visibilities and delay power spectra are dependent on right ascension (RA) of the pointing as well as the frequency band. We defer expressing visibilities and delay power spectra explicitly as a function of pointing or center frequency (or redshift) to later sections where it becomes necessary.

@@

Considering that the dynamic range required to suppress foreground contamination in the {\it EoR window} may probably be even higher than that provided by a {\it Blackman-Harris} window function, we use a modified version given by the convolution of a {\it Blackman-Harris} window with itself:
\begin{align}\label{eqn:bhw2}
  W(f) &= W_\textrm{BH}(f) \ast W_\textrm{BH}(f).
\end{align}
In its Fourier domain, this window has a response obtained by squaring the {\it Blackman-Harris} window response and thus increases the dynamic range further in the power response by another 100 -- 120~dB while $\epsilon\approx 42\%$. This is an enormous gain in sidelobe characteristics for a negligible loss in over sensitivity relative to a simple {\it Blackman-Harris} weighting. This will ensure that the contamination resulting from spillover of foregrounds along the line-of-sight $k$-modes are limited by intrinsic spectral structures in the foregrounds and the beam patterns and not by sidelobes from spectral weighting. In this paper, we apply this modified spectral weighting function defined in equation~\ref{eqn:bhw2} whenever foregrounds are represented in the Fourier delay domain.

@@

\begin{figure*}[htb]
\centering
\subfloat[][EoR model 1, 150~MHz subband ]{\label{fig:21cmfast-fg-ratio-150}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_21cmfast_chrmbeam_sim_150.0_MHz_subband.eps}}
\subfloat[][EoR model 2, 150~MHz subband ]{\label{fig:lidz-fg-ratio-150}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_lidz_chrmbeam_sim_150.0_MHz_subband.eps}} \\
\subfloat[][EoR model 1, 170~MHz subband ]{\label{fig:21cmfast-fg-ratio-170}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_21cmfast_chrmbeam_sim_170.0_MHz_subband.eps}}
\subfloat[][EoR model 2, 170~MHz subband ]{\label{fig:lidz-fg-ratio-170}\includegraphics[width=0.45\linewidth]{eor_fg_ratio_llim_lidz_chrmbeam_sim_170.0_MHz_subband.eps}}
\caption{Lower limits on EoR sensitivity, $\rho(\boldsymbol{b},f_0,\alpha)$, in {\it foreground avoidance} modes for different HERA-19 baselines as a function of RA ($\alpha$) of pointing on the sky. Each panel corresponds to the EoR model and subband indicated. The baseline vector is specified by its length (in m) and orientation (in degrees) at the bottom and top of the $x$-axis respectively. The color scale is logarithmic. Even with worst-case sensitivity, clearly HERA will be able detect EoR drawn from independent models on all baselines for a large fraction of a full 24~hour sky transit cycle. The sensitivity is seen to increase with increasing baseline length. This is because the loss in sensitivity to diffuse foreground power with increase in antenna spacings is faster than in the case of the EoR signal.}
\label{fig:eor-fg-ratios}
\end{figure*}
